Ridgeplots 

Load in our libraries

install(tidyverse) # loads in all the tidyverse libraries
library(lubridate) # to make dealing with dates easier
library(ggridges) # the brand new ggjoy package

bugs <- as_data_frame(read.csv("insect_data.csv"))

```{r}
# load in libraries
library(tidyverse)  # loads in all the tidyverse libraries
library(lubridate)  # to make dealing with dates easier
library(ggridges)   # the brand new ggjoy package
```
```{r}
## Dataset: 
# I'm using a dataset that contains information about what bugs showed up in a light trap set up  # on the roof of the University of Copenhagen's Zoological Museum and when. 

# read in data & convert it to a tibble
bugs <- as_data_frame(read.csv("insect_data.csv"))

head(bugs)
```
```{r}
## Add a coulmn with the month of each observation

# mdy() tells the lubridate package what format our dates are in; i.e., "2003-07-15"
# month() says we only want the month from the date; i.e., 1-12
bugs$month  <- month(mdy(bugs$date1))
```
```{r}
# list of months for labelling graph
monthList <- c("Jan","Feb","Mar","April","May", "June","July","Aug","Sep","Oct","Nov","Dec")

# remap months from numbers (3:12) to words (March-December)
bugs$month <- plyr::mapvalues(bugs$month, levels(as.factor(bugs$month)), monthList[3:12])
```
```{r}
# plot the nubmer of bugs caught by month
ggplot(data = bugs, aes(x = month, y = individuals)) + geom_point() + 
  scale_x_discrete(limit=monthList)
```
Comment:
Most insects show up in the summer. (Denmark is in the Northern Hemisphere, so summer runs from June to September). 

```{r}
# get the total number of insects observed on each day (binning over years)
bugs$dayInYear <- yday(mdy(bugs$date1))

# joyplot of when insects were observed by order (this is a variable). Scale changeshow tall the peaks are
ggplot(data = bugs, aes(x = dayInYear, y = order)) + geom_density_ridges(scale = 0.9) + 
    theme_ridges()
```
Comment:
It looks like both orders (beetles & butterflies) tend to show up at roughly the same time
```{r}
# how events have shifted over time

# ridgeplot of dates on which insects were observed by year of observation
ggplot(data = bugs, aes(x = dayInYear, y = as.factor(year))) + 
    geom_density_ridges(scale = 0.9) + theme_ridges()
```
Comment: Maybe a little bit of shift.
```{r}
# look at the variance 
varianceByYear <- bugs %>% group_by(year) %>% summarise(variance = sd(dayInYear))

# plot variance by year
ggplot(varianceByYear, aes(year, variance)) + geom_line() + 
   geom_smooth(method='lm') # this function adds the fitted line (w/ confidence interval)
```

Comment:
It looks like there been increasing variance in what day of the year insects are observed in this light trap.










